---
title: "heatmap_18S"
author: "Johanna"
date: "2025-03-18"
output: html_document
---

## Libraries
Load following libraries in your Rstudio: 
```{r}
library(phyloseq)
library(dplyr)
library(tidyr)
library(FSA)
library(ggpubr)
library(stats)
library(multcompView)
library(writexl)
library(pheatmap)
library(RColorBrewer)
library(grDevices)
library(ggplot2)
library(vegan)
library(tibble)
```

## **1. Clean Data Filter (OPTIONAL) **
### **1.1 Remove Zero-Abundance Taxa and Samples**
```{r}
PS <- phyloseq_overall_final  #here you we use the phyloseq-object we generated in "02d_filter_phyloseq_create_subgroups18S.Rmd"
total_sequences <- sum(otu_table(PS))

# Remove taxa with zero abundance and keep those that have over 100 reads
taxa_filtered <- prune_taxa(taxa_sums(PS) > 0, PS)
taxa_filtered <- prune_taxa(taxa_sums(PS) >= 100, taxa_filtered)

# Remove samples with zero abundance 
PS_filt <- prune_samples(sample_sums(taxa_filtered) > 0, taxa_filtered)
```

### **1.2 Filter Based on Relative Abundance**
Keep the sequences that are abundant in 0.1% of the samples
```{r}
genus_abundance <- taxa_sums(PS_filt) / total_sequences
keep_genera <- genus_abundance > 0.0001
PS_filter <- prune_taxa(keep_genera, PS_filt)
```
---

## **2. Taxonomic Aggregation to Genus level  and Normalization**

### **2.1 Aggregate at Genus Level**
```{r}
glom <- tax_glom(PS_filter, taxrank = "Genus")
ps.melt <- psmelt(glom)
```

### **2.2 Normalize Abundances and Compute Mean Scaled Abundance**
I want to have a dataframe, that calculated for each Genus a scaled abundance. So that i can peacefully compare the different locations within one genus. 
```{r}
genus_mean_abundance <- ps.melt %>%
  group_by(Genus) %>%
  summarise(Mean_Abundance = mean(Abundance, na.rm = TRUE))

df_normalized <- ps.melt %>%
  left_join(genus_mean_abundance, by = "Genus") %>%
  mutate(Scaled_Abundance = Abundance / Mean_Abundance)


mean_scaled_location <- df_normalized %>%
  group_by(location, Genus) %>%  # Gruppieren nach Standort und Genus
  summarise(Mean_Scaled_Abundance = mean(Scaled_Abundance, na.rm = TRUE), .groups = "drop") %>%
  mutate(Mean_Scaled_Abundance = round(Mean_Scaled_Abundance, 3))  

mean_scaled_location <- mean_scaled_location %>%
  mutate(Log_Mean_Scaled_Abundance = log10(Mean_Scaled_Abundance))

mean_scaled_location <- mean_scaled_location %>%
  group_by(Genus, location) %>%
  summarise(Log_Mean_Scaled_Abundance = mean(Log_Mean_Scaled_Abundance, na.rm = TRUE), .groups = "drop")

```
---
## 3. Generating Matrix for Heatmap

```{r}
log_scaled_abundance_matrix <- mean_scaled_location %>%
  pivot_wider(names_from = location, values_from = Log_Mean_Scaled_Abundance) %>%
  column_to_rownames(var = "Genus") %>%
  as.matrix()

# replace -Inf with the smallest finite value in the matrix to avoid distortion 

log_scaled_abundance_matrix[is.infinite(log_scaled_abundance_matrix)]  <- min(log_scaled_abundance_matrix[is.finite(log_scaled_abundance_matrix)])
```
Here, I am defining the colors and that 0 is white! (OPTIONAL)
```{r}
# define the value range of the heatmap
min_val <- min(log_scaled_abundance_matrix, na.rm = TRUE)
max_val <- max(log_scaled_abundance_matrix, na.rm = TRUE)
center_val <- 0 

#create color area
breaks <- c(seq(min_val, center_val, length.out = 50), 
            seq(center_val, max_val, length.out = 51)[-1])  

heatmap_colors <- colorRampPalette(c("#2166AC", "#F7F7F7", "#B2182B"))(100)
```

## **4. Statistical Analysis**
I am comparing here first Kruskal-Wallis Test for each genus and search for significant differences. Afterwards I am comparing the different
  locations using the Dunn-Posthoc test to test the locations with one-another. Plus I also calculate the obverall Revaltive abundance of the Genera. So we know what we see often in the samples. 
### **4.1 Kruskal-Wallis Test**
```{r}
kruskal_results <- df_normalized %>%
  group_by(Genus) %>%
  summarise(
  p_value = kruskal.test(Scaled_Abundance ~ location)$p.value
  ) %>%
  mutate(p_adjusted = p.adjust(p_value, method = "BH"))  # Multiple Testkorrektur mit Benjamini-Hochberg

# **Genera mit signifikanten Unterschieden auswählen (p < 0.05)**
significant_genera <- kruskal_results %>%
  filter(p_adjusted < 0.05) %>%
  pull(Genus)

```
### **4.2 Dunn’s Post Hoc Test**
And safe the significant values (and also the not-significant ones) 
```{r}
dunn_results <- df_normalized %>%
  filter(Genus %in% significant_genera) %>%
  group_by(Genus) %>%
  do({
    test <- dunnTest(Scaled_Abundance ~ location, data = ., method = "bh")
    data.frame(Genus = unique(.$Genus), test$res)
  }) %>%
  ungroup()

write_xlsx(dunn_results, "~/Desktop/SCHISDIMA/18S_Mada_Germany/dunnresults.xlsx")

dunn_significant <- dunn_results %>%
  filter(P.adj < 0.05)

write_xlsx(dunn_significant, "~/Desktop/SCHISDIMA/18S_Mada_Germany/dunnresults_18S_sig.xlsx")
```

### 4.3. Calculate Overall Relative Abundance regarding the genera
```{r}

ps.rel <- transform_sample_counts(PS_filter, function(x) x / sum(x))  # Konvertiere zu relativen Abundanzen

glom <- tax_glom(ps.rel, taxrank = "Genus")  # Gruppiere auf Genus-Ebene
ps.melt <- psmelt(glom)  # Konvertiere phyloseq-Objekt zu DataFrame

# Entferne NA-Werte in der Genus-Spalte
ps.melt <- ps.melt %>%
  filter(!is.na(Genus))

ps.melt_sum_location <- ps.melt %>%
  group_by(location, Genus) %>%
  summarise(mean_abundance = mean(Abundance), .groups = "drop")  # Berechne den Mittelwert pro Genus & Location


abundance_wide <- ps.melt_sum_location %>%
  pivot_wider(names_from = location, values_from = mean_abundance, values_fill = 0) %>%  # Fehlende Werte mit 0 füllen
  column_to_rownames("Genus")

abundance_matrix <- as.matrix(abundance_wide)  # Konvertiere in eine Matrix
overall_relative_abundance <- rowMeans(abundance_matrix, na.rm = TRUE)  # Mittelwert pro Genus über alle Locations hinweg
log10_abundance <- log10(overall_relative_abundance + 1)  # Log-Transformation (1 hinzugefügt, um log(0) zu vermeiden)
```

## 5. Create the Heatmap
### 5.1. Colors/Annotations
Perform annotation to rows/columns 
```{r}
annotation_row <- data.frame( 
  Significant_Differenced_in_abundance = kruskal_results$p_adjusted ,
  Overall_RelativeAbundance = log10_abundance,
  # Füge berechnete Log10-Abundanz hinzu
  Classification = c(
    "Fungi", "Protist", "Helminth", "Fungi", "Fungi", "Fungi", "Fungi", "Protist", 
    "Protist", "Fungi", "Fungi", "Fungi", "Protist", "Fungi", "Fungi", "Helminth", 
    "Fungi", "Protist", "Fungi", "Fungi", "Helminth", "Fungi", "Fungi", "Fungi"),
  row.names = rownames(abundance_matrix) # Setze Genus-Namen als Zeilennamen
)
                                  
annotation_col <- data.frame(
   Urbanization = c("rural", "urban", "semi-urban", "urban"),
   row.names = colnames(log_scaled_abundance_matrix)  
 )

adjusted_colors <- function(color, alpha = 0.6) {
  adjustcolor(color, alpha.f = alpha)
}

# Neue Farben für Classification mit "Alpha"-Effekt
ann_colors <- list(
  Urbanization = c("rural" = "thistle1",   
                   "semi-urban" = "thistle3", 
                   "urban" = "thistle4"),  
  Classification = c("Fungi" = adjusted_colors("#ff4500cc", 0.6),   
                     "Helminth" = adjusted_colors("#7570b3e5", 0.6), 
                     "Protist" = adjusted_colors("#66BD63", 0.6)),  # 40% Transparenz
  Overall_RelativeAbundance = colorRampPalette(c("#F7F7F7", "#DABC13"))(100),
  sign_differences_KrutisWallis = c("***" = "#A0A0A0",   # Dark Grey (Strong Significan
                         "**"  = "#C0C0C0",   # Medium-Dark Grey
                         "*"   = "#E5E5E5",   # Light Grey
                         "ns"  = "white")   # Very Light Grey (Almost White)
)                             
```

### 5.3. Draw the Heatmap
```{r}
pheatmap(
  log_scaled_abundance_matrix ,  
  scale = "none",  
  clustering_distance_cols = "manhattan",  
  clustering_method = "ward.D",
  show_rownames = TRUE,
  show_colnames = TRUE,
  border_color = "white",
  main = "Heatmap of Log-Scaled Abundances per Genus & Location",
  fontsize_row = 12,
  fontsize_col = 12,
  cellwidth = 50,
  cellheight = 15,
  color = heatmap_colors,
  annotation_col = annotation_col,  
  annotation_row = annotation_row,  
  annotation_colors = ann_colors, 
  breaks = breaks, 
  fontsize = 12 
)
```

### 5.4. Horizontal Heatmap (OPTIONAL) 
#transpose all the matrices
log_scaled_abundance_matrix_t <- t(log_scaled_abundance_matrix)  # Transponiert die Matrix
annotation_row_t <- annotation_col  # Locations werden jetzt Reihen-Annotationen
annotation_col_t <- annotation_row  # Genera werden jetzt Spalten-Annotationen
ann_colors_t <- list(
  Urbanization = ann_colors$Urbanization,  
  Classification = ann_colors$Classification,
  Overall_RelativeAbundance = ann_colors$Overall_RelativeAbundance,
  sign_differences_KrutisWallis = ann_colors$sign_differences_KrutisWallis
)

# horizontal Heatmap (transposed)
pheatmap(
  log_scaled_abundance_matrix_t,  
  scale = "none",  
  clustering_distance_rows = "manhattan",  # Jetzt Zeilen clustern (ehemals Spalten)
  clustering_method = "ward.D",
  show_rownames = TRUE,  # Zeilen anzeigen (jetzt Locations)
  show_colnames = TRUE,  # Spalten anzeigen (jetzt Genera)
  border_color = "white",
  main = "Relative abundance of genera per location",
  fontsize_row = 12,
  fontsize_col = 12,
  cellwidth = 20,
  cellheight = 20,
  color = heatmap_colors,
  annotation_col = annotation_col_t,# Jetzt Spalten-Annotation für Genera
  annotation_row = annotation_row_t,  # Jetzt Zeilen-Annotation für Locations
  annotation_colors = ann_colors_t,  # Gleiche Farbzuordnung beibehalten
  breaks = breaks,  
  angle_col = "45", 
  fontsize = 12 
)
```
                                  





