---
title: "binary_values_eukaryotes"
author: "Johanna"
date: "2025-03-04"
output: html_document
---


## Libraries
```{r}
library(phyloseq) # Because we are going to work with microbioe data using phyloseq object
library(vegan) # Many ecology related functions
library(tidyverse) # Use of pipe and lots of functions for a better R coding
library(microbiome) # some useful functions
library(DESeq2)# Test with negative binomial models
library(dplyr) 
library(dada2)
library(magrittr)
library(readxl)
library(Biostrings)
library(ggplot2)
library(ggpubr)
library(patchwork)
````

## Binary Values Eukaryotic Reads
In here, I want to analyse the data I generated. To have an overview about who has certain reads and 
calculate the abundance of certain eukaryotic taxa according to the location.
I am using the phyloseq-object "phyloseq_overall_final" I generates in "02d_filter_phyloseq_create_subgroups.Rmd"

```{r}
sample_metadata <- sample_data(phyloseq_overall_final)

# Define the locations to check
locations <- c("Kiel", "FA_Andina", "TO_Tsiroanomandidy", "MA_Mahajanga_Ankazomborona")
```

### Location
Investigate the metrics for the locations
```{r}
# Define the locations to check
locations <- c("Kiel", "FA_Andina", "TO_Tsiroanomandidy", "MA_Mahajanga_Ankazomborona")

# Count the number of samples for each location
location_counts <- sapply(locations, function(loc) sum(sample_metadata$location == loc, na.rm = TRUE))

# Print the results
location_counts
```
### Sex
Investigate the metrics for the sex of the study participants
```{r}
sex_counts <- lapply(locations, function(loc) {
  data_subset <- subset(sample_metadata, location == loc)
  list(
    female = sum(data_subset$sex== "female", na.rm = TRUE),
    male = sum(data_subset$sex == "male", na.rm = TRUE)
  )
})

overall_sex_counts <- list(
  female = sum(sample_metadata$sex == "female", na.rm = TRUE),
  male = sum(sample_metadata$sex == "male", na.rm = TRUE)
```
### Age
Investigate the metrics for the age of the study participants
```{r}
age_summary <- lapply(locations, function(loc) {
  data_subset <- subset(sample_metadata, location == loc & !is.na(age))
  
  if (nrow(data_subset) > 0) {
    list(
      Mean_Age = mean(data_subset$age, na.rm = TRUE),
      Age_Range = paste0(min(data_subset$age, na.rm = TRUE), " - ", max(data_subset$age, na.rm = TRUE))
    )
  } else {
    list(
      Mean_Age = NA,
      Age_Range = "No Data"
    )
  }
})


overall_age_summary <- list(
  Mean_Age = mean(sample_metadata$age, na.rm = TRUE),
  Age_Range = paste0(min(sample_metadata$age, na.rm = TRUE), " - ", max(sample_metadata$age, na.rm = TRUE))
)
```
Now I wanna know regarding some eukaryotic taxa of interest how many individuals had reads.
I am starting with different fungi, food-associated fungi, continute with helminths and finish with protozoa.
  
```{r}
taxa_of_interest_fungi <- c("Saccharomyces", "Debaryomyces", "Mucor", "Penicillium", "Starmerella", "Galactomyces", "Issatchenkia", "Candida", "Kazachstania", "Hanseniaspora", "Rhizopus", "Microascus", "Aspergillus", "Torulaspora", "Pichia", "Cryptococcus", "Lodderomyces", "Cladosporium", "Trichosporon", "Malassezia", "Sordaria")
taxa_of_interest_food_fungi <- c("Saccharomyces", "Debaryomyces",  "Mucor", "Penicillium",   "Starmerella",   "Galactomyces",  "Issatchenkia",  "Candida", "Kazachstania",  "Hanseniaspora", "Rhizopus",  "Microascus",   "Aspergillus",   "Torulaspora",   "Pichia")
taxa_of_interest_helminths <- c("Schistosoma", "Necator", "Enterobius", "Ascaris", "Trichuris", "Ancylostoma", "Strongyloides", "Bothriocephalus")
taxa_of_interest_protozoa <- c("Blastocystis", "Entamoeba", "Dientamoeba", "Pentatrichomonas", "Leidyana1")
```
  
Calculate the numbers of participants carrying reads:
```{r}
genus_interested <- subset_taxa(phyloseq_overall_final, Genus %in% taxa_of_interest_fungi) #here you type in one of the 4 different taxa_of_interest-Lists
samples_interested <- sample_names(genus_interested)[sample_sums(genus_interested) > 0]

phyloseq_with_interested_samples <- prune_samples(samples_interested, phyloseq_overall_final)
sample_metadata <- sample_data(phyloseq_with_interested_samples)

locations <- c("Kiel", "FA_Andina", "TO_Tsiroanomandidy", "MA_Mahajanga_Ankazomborona")

# Count the number of samples for each location
location_counts <- sapply(locations, function(loc) sum(sample_metadata$location == loc, na.rm = TRUE))
# Print the results
location_counts
```

Now check for the different Genera you wanna look at: 
```{r}
genus_ofinterest <- subset_taxa(phyloseq_overall_final, Genus %in% "Schistosoma") #Here you add the Genus you wanna have the information about! 

samples_ofinterest <- sample_names(genus_ofinterest)[sample_sums(genus_ofinterest) > 0]

# Create a new phyloseq object containing only these samples
phyloseq_with_samples_ofinterest <- prune_samples(samples_ofinterest, phyloseq_overall_final)

sample_metadata <- sample_data(phyloseq_with_samples_ofinterest)

# Define the locations to check
locations <- c("Kiel", "FA_Andina", "TO_Tsiroanomandidy", "MA_Mahajanga_Ankazomborona")

# Count the number of samples for each location
location_counts <- sapply(locations, function(loc) sum(sample_metadata$location == loc, na.rm = TRUE))

# Print the results
location_counts
```

                          

                          

                          


