---
title: "alpha_diversity_16S"
author: "Johanna"
date: "2025-02-27"
output: html_document
---

# Preparations

## Set global options

```{r style, echo = FALSE, results="asis", cache=FALSE, message=FALSE}
# v.05

# This piece of code is necessary if running Rmd on vscode,
# which does not read the header
if (!exists("params")) {
  params <- list(fig_path = "figures/", d_out = "./")
}

# Set knit global options
library("knitr")
options(digits = 2, width = 80)
golden_ratio <- (1 + sqrt(5)) / 2
opts_chunk$set(
  # General chunk configuration
  cache = FALSE,
  warning = FALSE, # Make warning FALSE, unless explicitly turn it on
  echo = TRUE,
  tidy = FALSE,
  include = TRUE,
  # Figure configurations
  fig.path = params$fig_path,
  dev = c("png"),
  fig.height = 10,
  fig.width = 10 * golden_ratio,
  fig.wide = TRUE,
  out.width = "150%", # Make figures appear larger and fit to the page width
  dpi = 300,
  comment = "  "
)

# Set seed for reproducibility
set.seed(123)

# Set plot options
library("ggplot2")
library("ggthemes")
theme_set(theme_bw(base_size = 20))


#  Set color blind friendly pallet as default for discrete colors
# Source of palette: https://www.sthda.com/english/articles/32-r-graphics-essentials/125-ggplot-cheat-sheet-for-great-customization #nolint
cbp1 <- c(
  "#E69F00",
  "#56B4E9",
  "#009E73",
  "#F0E442",
  "#0072B2",
  "#D55E00",
  "#CC79A7",
  "#999999"
)

dynamic_cbp1_colors <- function(n) {
  colorRampPalette(cbp1)(n)
}

# Define custom scales using the dynamic color generator
scale_fill_custom <- function(...) {
  discrete_scale(
    aesthetics = "fill",
    scale_name = "custom",
    palette = dynamic_cbp1_colors,
    ...
  )
}
scale_colour_custom <- function(...) {
  discrete_scale(
    aesthetics = "colour",
    scale_name = "custom",
    palette = dynamic_cbp1_colors, ...
  )
}

# Set the custom scales as default
options(
  ggplot2.discrete.fill = scale_fill_custom,
  ggplot2.discrete.colour = scale_colour_custom
)

# Set viridis as default for scale continuous
options(
  ggplot2.continuous.fill = scale_fill_viridis_c,
  ggplot2.continuous.colour = scale_colour_viridis_c
)

#' Display Data Frame with Formatted Numeric Columns
#'
#' This function formats a data frame by rounding numeric columns and
#' displaying it either as a static table using `knitr::kable` or as an
#' interactive table using `DT::datatable`.
#'
#' @param data A data frame to be formatted and displayed.
#' @param type A character string specifying the display type: "static"
#' for a static table using `knitr::kable`,
#' "interactive" for an interactive table using `DT::datatable`.
#' Default is "static".
#' @param n_round An integer specifying the number of decimal places to round
#'  numeric columns. Default is 3.
#'
#' @return Formatted data frame displayed as specified.
#' @examples
#' \dontrun{
#' show_df(data, type = "static")
#' show_df(data, type = "interactive", n_round = 2)
#' }
show_df <- function(data, type = "static", n_round = 3) {
  data <- data %>% mutate_if(is.numeric, \(x) round(x, n_round))
  if (type == "static") {
    data <- knitr::kable(data, row.names = FALSE)
  }
  if (type == "interactive") {
    data <- DT::datatable(data,
      rownames = FALSE, extensions = "FixedColumns",
      options = list(scrollX = TRUE, fixedColumns = list(leftColumns = 1))
    )
  }
  data
}
# Set output directory
d_out <- params$d_out
rm(params)
```

## Libraries
```{r}
library(phyloseq) # Because we are going to work with microbioe data using phyloseq object
library(vegan) # Many ecology related functions
library(tidyverse) # Use of pipe and lots of functions for a better R coding
library(microbiome) # some useful functions
library(DESeq2)# Test with negative binomial models
library(dplyr) 
library(dada2)
library(magrittr)
library(readxl)
library(Biostrings)
library(ggplot2)
library(ggpubr)
library(patchwork)
````

## Analyse Alpha Diversity 
For total 16S dataset. I choosed to calculate the Shannon-Diversity, but you can also do so with Chao1, Observed or any other AlphaDiversity Index. 
```{r}
rich.Shannon <- 
  pruned_bacteria_eukaryotes %>%
  estimate_richness(measures = "Shannon") %>% 
  rownames_to_column("pid")

Alpha.Diversity <- cbind(sample_data(pruned_bacteria_eukaryotes), rich.Shannon)

Alpha.Diversity$location[Alpha.Diversity$location == "Mahajanga"] <- "Ankazomborona" #Its Mahajanga_Ankazomborona, but I use Ankazomborona, cause this is more accurate for the desribtion of the location
Alpha.Diversity$location <- factor(Alpha.Diversity$location,  levels = c( "Andina", "Ankazomborona", "Tsiroanomandidy","Kiel")) #ggplot needs a factor 
  
Alpha.Diversity <- Alpha.Diversity %>%
                    select(-1)          #remove the dual "PID" column due to the cbind before 
```
## Plot Alpha Diversity
and safe it directs as a PDF! 
```{r}
pdf("AlphaDiversity_16S.pdf", width = 8, height = 8) 

ggplot(Alpha.Diversity, aes(x = location, y = Shannon)) +
  geom_jitter(aes(color = factor(location)), width = 0.3, height = 0, alpha = 1, size = 2)+
  geom_boxplot(aes(group = location), width = 0.8, fill = NA, color = "black", size = 1.2) +
  scale_color_manual(values = c("Ankazomborona" = "thistle3","Andina" = "thistle2", "Kiel" = "#706c70", "Tsiroanomandidy" = "thistle4"),
    name = "Location")+
  labs(x = "Location", y = "Shannon") + 
  theme_minimal()+
  theme(panel.grid.major.y = element_line(color = "grey85", linetype = "dashed", size = 0.5),  # Dezente horizontale Linien
        panel.grid.major.x = element_blank(),  # Vertikale Linien entfernen
        panel.background = element_rect(fill = "white", color = NA),
        legend.position = "right",
        legend.title = element_text(size = 12, face = "bold"),
        legend.text = element_text(size = 12),
        axis.text = element_text(size = 14),        
        axis.title = element_text(size = 16, face = "bold"),
        axis.text.x = element_text(angle = 45, hjust = 1)) + 
  stat_compare_means(comparisons = list(
                                        c("Tsiroanomandidy", "Ankazomborona"), 
                                        c("Ankazomborona", "Andina"), 
                                        c("Kiel", "Ankazomborona")), 
                       method = "wilcox.test", label = "p.signif", 
                       label.y = c(5.1, 5.5, 5.7, 5.9, 6.2, 6.4), vjust = 0.3)
dev.off()
```

## Calculate the Numbers
We wanna have the means of the Shannon-Diversities as well as the p-values of the Wilcoxon-Rank-Sum-Test (bonferroni - adjustment. there is not a real difference, regarding the results, also if you do BH testing!)
```
summary_stats <- Alpha.Diversity %>%
  group_by(location) %>%
  summarise(
    Median_Shannon = median(Shannon, na.rm = TRUE),
    Min_Shannon = min(Shannon, na.rm = TRUE),
    Max_Shannon = max(Shannon, na.rm = TRUE)
  )
# Ergebnis anzeigen
print(summary_stats)
``
location        Median_Shannon Min_Shannon Max_Shannon
  <fct>                    <dbl>       <dbl>       <dbl>
1 Andina                    3.90       1.98         5.17
2 Ankazomborona             4.44       2.74         5.12
3 Tsiroanomandidy           3.97       0.954        4.90
4 Kiel                      3.99       0.506        5.58


```{r}
wilcox_results <- pairwise.wilcox.test(
  Alpha.Diversity$Shannon, 
  Alpha.Diversity$location, 
  p.adjust.method = "bonferroni" # oder "BH"
)

# Ergebnisse anzeigen
print(wilcox_results)
````
	Pairwise comparisons using Wilcoxon rank sum test with continuity correction 

data:  Alpha.Diversity$Shannon and Alpha.Diversity$location 

                Andina Ankazomborona Tsiroanomandidy
Ankazomborona   <2e-16 -             -              
Tsiroanomandidy 1.00   <2e-16        -              
Kiel            0.46   <2e-16        1.00           

P value adjustment method: bonferroni 

