---
title: "beta_diversity_16S"
author: "Johanna"
date: "2025-03-03"
output: html_document
---

# Preparations

## Set global options

```{r style, echo = FALSE, results="asis", cache=FALSE, message=FALSE}
# v.05

# This piece of code is necessary if running Rmd on vscode,
# which does not read the header
if (!exists("params")) {
  params <- list(fig_path = "figures/", d_out = "./")
}

# Set knit global options
library("knitr")
options(digits = 2, width = 80)
golden_ratio <- (1 + sqrt(5)) / 2
opts_chunk$set(
  # General chunk configuration
  cache = FALSE,
  warning = FALSE, # Make warning FALSE, unless explicitly turn it on
  echo = TRUE,
  tidy = FALSE,
  include = TRUE,
  # Figure configurations
  fig.path = params$fig_path,
  dev = c("png"),
  fig.height = 10,
  fig.width = 10 * golden_ratio,
  fig.wide = TRUE,
  out.width = "150%", # Make figures appear larger and fit to the page width
  dpi = 300,
  comment = "  "
)

# Set seed for reproducibility
set.seed(123)

# Set plot options
library("ggplot2")
library("ggthemes")
theme_set(theme_bw(base_size = 20))


#  Set color blind friendly pallet as default for discrete colors
# Source of palette: https://www.sthda.com/english/articles/32-r-graphics-essentials/125-ggplot-cheat-sheet-for-great-customization #nolint
cbp1 <- c(
  "#E69F00",
  "#56B4E9",
  "#009E73",
  "#F0E442",
  "#0072B2",
  "#D55E00",
  "#CC79A7",
  "#999999"
)

dynamic_cbp1_colors <- function(n) {
  colorRampPalette(cbp1)(n)
}

# Define custom scales using the dynamic color generator
scale_fill_custom <- function(...) {
  discrete_scale(
    aesthetics = "fill",
    scale_name = "custom",
    palette = dynamic_cbp1_colors,
    ...
  )
}
scale_colour_custom <- function(...) {
  discrete_scale(
    aesthetics = "colour",
    scale_name = "custom",
    palette = dynamic_cbp1_colors, ...
  )
}

# Set the custom scales as default
options(
  ggplot2.discrete.fill = scale_fill_custom,
  ggplot2.discrete.colour = scale_colour_custom
)

# Set viridis as default for scale continuous
options(
  ggplot2.continuous.fill = scale_fill_viridis_c,
  ggplot2.continuous.colour = scale_colour_viridis_c
)

#' Display Data Frame with Formatted Numeric Columns
#'
#' This function formats a data frame by rounding numeric columns and
#' displaying it either as a static table using `knitr::kable` or as an
#' interactive table using `DT::datatable`.
#'
#' @param data A data frame to be formatted and displayed.
#' @param type A character string specifying the display type: "static"
#' for a static table using `knitr::kable`,
#' "interactive" for an interactive table using `DT::datatable`.
#' Default is "static".
#' @param n_round An integer specifying the number of decimal places to round
#'  numeric columns. Default is 3.
#'
#' @return Formatted data frame displayed as specified.
#' @examples
#' \dontrun{
#' show_df(data, type = "static")
#' show_df(data, type = "interactive", n_round = 2)
#' }
show_df <- function(data, type = "static", n_round = 3) {
  data <- data %>% mutate_if(is.numeric, \(x) round(x, n_round))
  if (type == "static") {
    data <- knitr::kable(data, row.names = FALSE)
  }
  if (type == "interactive") {
    data <- DT::datatable(data,
      rownames = FALSE, extensions = "FixedColumns",
      options = list(scrollX = TRUE, fixedColumns = list(leftColumns = 1))
    )
  }
  data
}
# Set output directory
d_out <- params$d_out
rm(params)
```

## Libraries
```{r}
library(phyloseq) # Because we are going to work with microbioe data using phyloseq object
library(vegan) # Many ecology related functions
library(tidyverse) # Use of pipe and lots of functions for a better R coding
library(microbiome) # some useful functions
library(DESeq2)# Test with negative binomial models
library(dplyr) 
library(dada2)
library(magrittr)
library(readxl)
library(Biostrings)
library(ggplot2)
library(ggpubr)
library(patchwork)
````

## Analyse Beta Diversity 
We analyse how diverse the samples are compared to each other with analysing the beta diversity using non-metric multidimensional scaling (NMDS) and the Bray-Curtis dissimilarity index. 
Plot the beta diversity first and then perform the statistical test. 
Here, we use the phyloseq-object "pruned_bacteria_eukaryotes", that we generated in 02d_filter_phyloseq_16S.Rmd. 


````{r}
sample_data(pruned_bacteria_eukaryotes)$location[sample_data(pruned_bacteria_eukaryotes)$location == "Mahajanga"] <- "Ankazomborona" #adjust again the location names here 

set.seed(123)
ps.ord <-
  pruned_bacteria_eukaryotes %>% 
  ordinate("NMDS", "bray")




