---
title: "qPCR_Analysis"
author: "Johanna Saalfrank"
date: "2024-12-11"
output: html_document
---

## Overview

qPCR Analysis of total 50 samples from Kiel and Mahajanga samples was performed. 
R Studio is used to generate the concentration plot. 
The concentration was calculated by using the standard curve from the run performed on Dec. 09, 2024. 
The concentration is measured in pg /sample. The protocol is desribed in the Paper. 

Following libraries will be used:

```{r}
library(readxl)
library(ggplot2)
library(ggpubr)
```
Load the excel-table (from LightCycler (R) 480 ) including the S.cerevisiae concentrations for all samples, remove two with a faulty concentration of very hight values (S-0150-K0 and S-0029-K0) 

```{r}
ggplot(qPCR_Analysis_filtered, aes(x = location, y = `conc in pg`)) +
    geom_boxplot(aes(group = location), width = 0.5, fill = NA, color = "black", size = 1, outlier.shape = NA) +
    geom_jitter(aes(color = location), width = 0.2, height = 0, alpha = 0.5, size = 3) + 
    stat_summary(
        fun = median, 
        geom = "point", 
        shape = 18, 
        size = 3, 
        color = "black"
    ) +                                     # Add median as a balck point not visible 
    stat_summary(
        fun = median,
        geom = "text",
        aes(label = sprintf("%.3f", ..y..)),
        vjust = -0.5,
        color = "black",
        size = 5
    ) +                                     # Add median value as text above the point
    stat_compare_means(
        method = "wilcox.test",                         # Wilcoxon test for medians
        label = "p.format",                             # Display p-value in a formatted style
        comparisons = list(c("germany", "madagascar")), # Compare these two groups
        label.y = 0.4                                   # Position of the p-value label on the y-axis
    ) +                                     # Add labels
    labs(
        x = "Location",
        y = "Absolute concentration of S.cerevisiae (in pg/sample)",
        color = "Location"                              
    ) +
    scale_y_continuous(limits = c(0, 1), oob = scales::squish)+
    scale_color_manual(
        values = c(
            "germany" = "#FFC107",
            "madagascar" = "#004D40"
        )
    ) +
    theme(
        panel.grid = element_blank(),
        panel.background = element_rect(fill = "white", color = NA),
        legend.position = "none",
        axis.text = element_text(size = 14),
        axis.title = element_text(size = 16, face = "bold")
    ) 

ggsave("F1_qPCR.pdf", width = 8.86 , height = 9.40, units "in")     #Save the Plot

``` 

