---
title: "qPCR_Analysis"
author: "Johanna Saalfrank"
date: "2024-12-11"
output: html_document
---

## Overview

qPCR Analysis of total 50 samples from Kiel and Mahajanga samples was performed. 
R Studio is used to generate the concentration plot. 
The concentration was calculated by using the standard curve from the run performed on Dec. 09, 2024. 
The concentration is measured in pg. 

Following libraries will be used:

```{r}
library(readxl)
library(ggplot2)
library(ggpubr)
```
Load the excel-table including the S.cere concentrations for all samples, remove two with a faulty concentration of very hight values (S-0150-K0 and S-0029-K0) 

```{r}
qPCR_Analysis <- read_excel("Desktop/qPCR Analysis.xlsx", + sheet = "combined")
#remove the outlayers
qPCR_Analysis_filtered <- qPCR_Analysis[qPCR_Analysis$Sample_name != "S-0150-K0", ]
qPCR_Analysis_filtered <- qPCR_Analysis_filtered[qPCR_Analysis_filtered$Sample_name != "S-0029-K0", ]
```
Plotting a boxplot, grouped by location of samples and showing the concentration of S. cere in fecal DNA, measured in pg. 
```{r}
ggplot(qPCR_Analysis_filtered, aes(x = location, y = `conc in pg`)) +
    geom_boxplot(aes(group = location), width = 0.5, fill = NA, color = "black", size = 1.2, outlier.shape = NA) +
    geom_jitter(aes(color = location), width = 0.3, height = 0, alpha = 0.3, size = 2) +   # Jitter for points colored by location
    stat_summary(
        fun = median, 
        geom = "point", 
        shape = 18, 
        size = 3, 
        color = "black"
    ) +      # median as a point
    stat_summary(
        fun = median,
        geom = "text",
        aes(label = sprintf("%.3f", ..y..)),
        vjust = -0.5,
        color = "black",
        size = 4
    ) +     # median value as text 
    stat_compare_means(
        method = "wilcox.test",                         # Wilcoxon test for medians
        label = "p.format",                             # Display p-value in a formatted style
        comparisons = list(c("germany", "madagascar")), # Compare these two groups
        label.y = 0.18                                  # Position of the p-value label on the y-axis
    ) +     # lables of the plot 
    labs(
        title = "Concentrations of Saccharomyces by Location",
        x = "Location",
        y = "Concentration of S.cere. (pg)",
        color = "Location" 
    ) +
    scale_y_continuous(limits = c(0, 1), oob = scales::squish)+ # accumulate the "higher" valued on the upper limit of the plot and set the limits of the plot 
    scale_color_manual(                                         # set the colors for the jitter points 
        values = c(
            "germany" = "#FFC107",
            "madagascar" = "#004D40"
        )
    ) +
    theme(                                                      # Make the plot pretty overall 
        panel.grid = element_blank(),
        panel.background = element_rect(fill = "white", color = NA),
        legend.position = "right",
        legend.title = element_text(size = 12, face = "bold"),
        legend.text = element_text(size = 10),
        axis.text = element_text(size = 14),
        axis.title = element_text(size = 16, face = "bold")
    )
```

This could be a add on to show how the distribution looks like. Bins are formed in here. First the concentration-results are binned 
in different values and in the second step a barplot is generated to compare the bins. 

```{r}
qPCR_Analysis_filtered$bins <- cut(                  #add a new column called bins
    qPCR_Analysis_filtered$`conc in pg`,             #add according to the value in `conc in pg` a label 
    breaks = c(0, 0.01, 0.02, 0.03, 0.04, 0.05, Inf),
    labels = c("0-0.01", "0.01-0.02", "0.02-0.03", "0.03-0.04", "0.04-0.05", ">0.05"),
    include.lowest = TRUE
)
#create a barplot 
ggplot(qPCR_Analysis_filtered, aes(x = bins, fill = location)) +
    geom_bar(position = "dodge", color = "black", size = 0.5) + # barplot with sperated bars per location
    labs(
        title = "Distribution of Saccharomyces Concentrations by Location",
        x = "Concentration Range (pg)",
        y = "Number of Samples",
        fill = "Location"
    ) +
    scale_fill_manual(                              # set the colors for the bars
        values = c(
            "germany" = "#FFC107",
            "madagascar" = "#004D40"
        )
    ) +
    theme_minimal() +                            # Make the plot pretty overall 
    theme(
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 14, face = "bold"),
        legend.title = element_text(size = 12, face = "bold"),
        legend.text = element_text(size = 10)
    )
```
